---
title: "Bellabeat Case Analysis"
author: "Leo Pham"
date: "2022-10-15"
output: html_document
---

## BELLABEAT CASE ANALYSIS

### INTRODUCTION

#### 1. Overview

Bellabeat, a high-tech manufacturer of health-focused smart products for women. Since 2013, Bellabeat has inspired women around the world to gain knowledge in improving one's health. Although Bellabeat has grown rapidly and quickly positioned itself as a large player in the global smart device market, Urška Sršen believes that analyzing smart fitness data can unlock new opportunities and insights for the company's business outlook.

#### 2. Products

-   Bellabeat App: The Bellabeat app provides users with health data related to their activity, sleep, stress, menstrual cycle, and mindfulness habits. This data can help users better understand their current habits and make healthy decisions. The Bellabeat app connects to their line of smart wellness products.
-   Leaf: Bellabeat's classic wellness tracker can be worn as a bracelet, necklace, or clip. The Leaf tracker connects to the Bellabeat app to track activity, sleep, and stress.

### ASK

#### 1. Business task

This project analyzes smart device usage data to gain insight into how people are already using their smart devices and give recommendations on how these trends can inform Bellabeat marketing strategy.

#### 2. Stakeholders

-   Urška Sršen: Bellabeat's co-founder and Chief Creative Officer.
-   Sando Mur: Mathematician and Bellabeat's cofounder, a key member of the Bellabeat executive team.
-   Bellabeat marketing analytics team: A team of data analysts responsible for collecting, analyzing, and reporting data that helps guide Bellabeat's marketing strategy. I am a junior data analyst who can help Bellabeat achieve its missions and goals.

### PREPARE

#### 1. Data description

-   Data source: FitBit Fitness Tracker Data [Link](https://www.kaggle.com/datasets/arashnic/fitbit).
-   Data organization: 18 CSV files organized in a long data format.
-   Sample size: 33 observations.
-   Data duration: 2016-03-12 to 2016-05-12.
-   Data credibility: 33 Fitbit users consented to the submission of personal tracker data, including minute-level output for physical activity, heart rate, and sleep monitoring. Since the data was collected by a third party, it is difficult to verify the reliability of it. Additionally, the data contains information from only 33 current Fitbit users, this could create a sampling bias.
-   Data license: CC0 Public Domain.

#### 2. Previewing data

Before importing the data, I use Excel spreadsheet to preview all the datasets and realize that those tables have covered a data scale from daily to minutes.

In this project, I will be focusing on analyzing daily data. Therefore, our research will be conducted on dailyActivity, dailyCalories, dailyIntensities, dailySteps, sleepDay, and weightLogInfo.

Additionally, dailyCalories, dailyIntensities, and dailySteps are the three subsets of dailyActivity so it is possible to remove these three tables out of my database. As a result, I am going to mainly perform data cleaning for there datasets dailyActivity, sleepDay, and weightLogInfo.

### PROCESS

I decide to use SQL for data cleaning because dailyActivity, sleepDay, and weightLogInfo are relational databases. The process will be starting with importing the csv files to SQL Management Studio and ending up with 2 files: Weight.csv and SleepAndActivity.csv.

Document of the cleaning process: cleaning.txt.

### ANALYZE

Data analyzing and data visualization are carried out on Rstudio.

Let's install and load some packages:

```{r}
install.packages("tidyverse")
installed.packages("tidyr")
install.packages("dplyr")
install.packages("ggplot2")
```

```{r}
library("tidyverse")
library("tidyr")
library("dplyr")
library("ggplot2")
```

After that, we import Weight and SleepAndActivity datasets.

```{r}
sleep_and_activity <- read.csv("E:/Downloads/SleepAndActivity.csv")
weight <- read.csv("E:/Downloads/Weight.csv")
```

```{r}
#Previewing SleepAndActivity
head(sleep_and_activity)
str(sleep_and_activity)

```

```{r}
#Previewing Weight dataset
head(weight)
str(weight)
```

Summary() is written down to gain statistical information about the dataset. I primarily use MEAN values to identify the findings.

```{r}
#Summarizing step, distance, and calories data
sleep_and_activity %>% select(total_steps,
                    total_distance,
                    calories) %>%
              summary()
```

Findings: The average step count is 8515 (somewhat active), according to Medical News Today taking 8,000 steps per day reduces 51% lower risk of dying by any cause. Individuals who walk up to 12,000 steps per day was associated with a 65% lower risk compared with taking 4,000 steps:

-   Sedentary: Less than 5,000 steps daily.

-   Low active: About 5,000 to 7,499 steps daily.

-   Somewhat active: About 7,500 to 9,999 steps daily.

-   Active: More than 10,000 steps daily.

-   Highly active: More than 12,500 steps daily.

```{r}
#Summarizing sleep data
sleep_and_activity %>% select(total_minutes_asleep,
                        total_time_in_bed) %>%
                  summary()
```

Findings: The average user sleeps about 6.99 hours per day which is almost in the recommended sleep hours for adults (7-9 hours) by sleepfoundation. However the average time in bed is higher than the average sleeping time per day around 40 minutes, which shows that participants have the tendency of staying on bed after they wake up.

```{r}
#Summarizing active minutes per category
sleep_and_activity %>% select(very_active_minutes,
                    fairly_active_minutes,
                    lightly_active_minutes,
                    sedentary_minutes) %>%
              summary()
```

Findings:

-   The average sedentary minutes is 712 minutes, which equates to 11.87 hours, which should be reduced.

-   Among lightly_active_minutes, fairly_active_minutes, and very_active_minutes; lightly_active_minutes has the highest mean of 216.5 minutes suggested that most participants engage in light activities.

```{r}
weight %>% select(weight_kg,
                  bmi) %>%
              summary()
```

Findings: The average BMI is 25.19, which is considered in the overweight category according to the standards from CDC:

-   If your BMI is less than 18.5, it falls within the underweight range.

-   If your BMI is 18.5 to \<25, it falls within the healthy weight range.

-   If your BMI is 25.0 to \<30, it falls within the overweight range.

-   If your BMI is 30.0 or higher, it falls within the obesity range.

### SHARE

Because there are only 8 distinct observations in the weight table, which is not significant enough to generate recommendations to our business goals. In this stage, we will mainly be construction visualizations for SleepAndActivity dataset.

```{r}
# User Types distribution
ggplot(data = sleep_and_activity) + geom_bar(mapping = aes(x = user_type, fill = user_type)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "User Types Distribution", x= "User Types", y = "Frequency", fill = "User Types")

```

Insights: More than 50% of the users are active or sedentary.

```{r}
ggplot(data = sleep_and_activity) + geom_bar(mapping = aes(x = sleep_quality, fill = sleep_quality)) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Sleep Quality Distribution", x= "Sleep Quality", y = "Frequency", fill = "Sleep Quality")

```

Insights: More than 50% of the users are oversleeping or having insufficient sleep.

```{r}
# Sleep quality and calories burned for each user type
ggplot(data = sleep_and_activity) + 
  geom_histogram(mapping = aes(x = calories, fill = sleep_quality, position = "identity")) +
  facet_wrap(~user_type) +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(title = "Sleep Quality and Calories Burned", y = "Frequency", x = "Calories Burned", fill = "Sleep Quality", subtitle = "For 5 different types of user")

```

Insights: Sedentary people with insufficient sleep tend to burn less calories than other user types

```{r}
# Total steps and Calories
ggplot(data = sleep_and_activity) +
  geom_point(mapping = aes(x=total_steps, y=calories)) + 
  geom_smooth(mapping = aes(x=total_steps, y=calories)) + 
  labs(title="Total Steps and Calories Burned", x= "Total Steps", y = "Calories Burned")

```

Insights: A positive correlation between the total steps and calories burned. This means that the more active the user is, the higher calories they need every day.

```{r}
# Total steps and tracker distance
ggplot(data = sleep_and_activity) +
  geom_point(mapping = aes(x=total_steps, y=tracker_distance)) + 
  geom_smooth(mapping = aes(x=total_steps, y=tracker_distance)) + 
  labs(title="Total Steps and Tracker Distance", x= "Total steps", y = "Tracker distance")

```

Insights: There is a strong positve correlation between the total steps and the tracker distance.

```{r}
# Very active minutes and calories burned
ggplot(data = sleep_and_activity) +
  geom_point(mapping = aes(x=very_active_minutes, y=calories)) + 
  geom_smooth(mapping = aes(x=very_active_minutes, y=calories)) + 
  labs(title="Very Active Minutes and Calories Burned", x= "Very Active Minutes", y = "Calories Burned")

```

Insights: The higher active minutes, the more calories burned.

```{r}
# Total minutes asleep and sedentary minutes
ggplot(data = sleep_and_activity) +
  geom_point(mapping = aes(x=sedentary_minutes, y=total_minutes_asleep)) + 
  geom_smooth(mapping = aes(x=sedentary_minutes, y=total_minutes_asleep)) + 
  labs(title = "Total Minutes Asleep and Sedentary Minutes", x= "Sedentary Minutes", y = "Amount of Sleep (minutes)")

```

Insights: There is a negative correlation between the total minutes asleep and sedentary minutes. It seems like users with higher sedentary minutes have lower sleep quality.

```{r}
# Total minutes asleep and very active minutes
ggplot(data = sleep_and_activity) +
  geom_point(mapping = aes(x=very_active_minutes, y=total_minutes_asleep)) + 
  geom_smooth(mapping = aes(x=very_active_minutes, y=total_minutes_asleep)) + 
  labs(title = "Total Minutes Asleep and Very Active Minutes", x= "Very Active Minutes", y = "Amount of Sleep (minutes)")

```

Insights: There is a positive correlation between the total minutes asleep and very active minutes.It seems like users with higher active minutes have better sleep quality.

### ACT
